4-4-netcdf

### 2022-02-07 1-intro.Rmd
small changes in read-through

Spatio-temporal data concerns variables in the spatial coordinates across time. Some variables are invariant to time and can be solely identified by its spatial location while other time varying variables need both the spatial and temporal identifier to identify. Manipulating data in these two levels is usually done separately and the third tidy data principal [@tidydata] recommends each type of observational unit to form a table. Another feature for a spatio-temporal data structure to have is that variables created from subsequent operations can be automatically added to the existing table with the same level. This is better than an output that index the variables with the identifiers since otherwise analysts will need to manually join this output with existing variables in the same level.

@@

Currently, available spatio-temporal data structure in R includes: \pkg{spacetime} [@spacetime], which proposes four space-time layouts: Full grid (STF), sparse grid(STS), irregular (STI), and trajectory (STT). The data structure it uses is based on \pkg{sp} [@sp] and \pkg{xts} [@xts], both of which has been replaced by more recent implementations. \pkg{spatstat} [@spatstat] implements a `ppp` class for point pattern data; and more recent, \pkg{stars} [@stars] implements a spatio-temporal array with the dplyr's data cube structure \pkg{cubelyr} [@cubelyr] as its back-end. While these implementations either store spatial and temporal variables all in a single table, hence duplicate the spatial variables for each temporal unit; or split them into two separate tables that has the problem of manually joining, mentioned in the previously. This creates a gap in the software development. The requirement for such a tool is important given the ubiquity of spatio-temporal vector data in the wild: the Ireland wind data from \pkg{gstat} is an classic example data that splits variables into its spatial (`wind.loc`) and temporal (`wind`) dimension; Bureau of Meteorology (BoM) provides climate observations widely applied in agriculture and ecology study.

This paper describes the implementation of a new spatio-temporal data structure: \pkg{cubble}. \pkg{cubble} implements a relational data structure that uses two forms to manage the switch between spatial and temporal dimension. With this structure, users can manipulate the spatial or temporal dimension separately, while leaves the linking of two dimensions to \pkg{cubble}. The software is available from the Comprehensive R Archive Network (CRAN) at [CRAN link].

The rest of the paper will be divided as follows: Section 2 presents the workflow of data manipulation in cubble. Section 3 explains how cubble deals with some advanced considerations  including data with hierarchical structure, data matching, and how cubble fits with existing static and interactive visualisation tools. Section 4 gives examples of the features introduced in the previous two sections with climate and hydrology data. An example on how cubble handle NetCDF data is also provided. Section 5 concludes the paper.

>>>>

Spatio-temporal data concerns variables in the spatial coordinates across time. Some variables are invariant to time and can be solely identified by its spatial location while other time varying variables need both the spatial and temporal identifiers to identify. Manipulating data in these two levels is usually done separately and the third tidy data principal [@tidydata] recommends each type of observational unit to form a table. Another feature for a spatio-temporal data structure to have is that variables created from subsequent operations can be automatically added to the existing table with the same level. This is better than an output that index the variables with the identifiers since otherwise analysts will need to manually join this output with existing variables in the same level.

@@

Currently, available spatio-temporal data structures in R includes: \pkg{spacetime} [@spacetime], which proposes four space-time layouts: Full grid (STF), sparse grid(STS), irregular (STI), and trajectory (STT). \pkg{spacetime} is built on \pkg{sp} [@sp] and \pkg{xts} [@xts], both of which have been replaced by more recent implementations. \pkg{spatstat} [@spatstat] implements a `ppp` class for point pattern data; and more recent, \pkg{stars} [@stars] implements a spatio-temporal for vector and raster spatio-temporal data.  While these implementations either store spatial and temporal variables all in a single table, hence duplicate the spatial variables at each temporal unit; or split them into two separate tables that analysts need to manually join the data constantly when doing the analysis. This creates a gap in the software development. The requirement for such a tool is important given the ubiquity of spatio-temporal vector data in the wild: climate observations from weather stations, air quality variables from air monitoring stations, or any variables that are measured at fixed location across time.

This paper describes the implementation of a new spatio-temporal data structure: \pkg{cubble}. \pkg{Cubble} implements a data structure that uses two forms to manage the switch between spatial and temporal dimension. With this structure, users can manipulate the spatial or temporal dimension separately, while leaves the linking of two dimensions to \pkg{cubble}. The software is available from the Comprehensive R Archive Network (CRAN) at [CRAN link].

The rest of the paper is divided as follows: Section 2 presents the workflow of data manipulation in cubble. Section 3 explains how cubble deals with some advanced considerations  including data with hierarchical structure, data matching, and how cubble fits with existing static and interactive visualisation tools. Section 4 gives examples of the features introduced in the previous two sections with climate and hydrology data. An example on how cubble handle NetCDF data is also provided. Section 5 concludes the paper.

### 2022-02-07 2-the-cubble-pkg.Rmd
small changes in read-through

Spatio-temporal data usually come in various forms and Figure \ref{fig:illu-input} shows four examples of this. These four examples show some common characters of the spatio-temporal data: 1) a spatial identifier, that uniquely identifies each site, 2) a temporal identifier, that prescribes the time stamp each site is recorded, 3) Coordinates, comprising of latitude and longitude, that locate each site on the map. These identifiers will be the building blocks for cubble. Spatio-temporal data can also be thought of as a data cube with three axes: id, time, and variable. Notice that in the time-variable face, time invariant variables share the same value across all the time stamps while time varying variables have values varying across time. This heterogeneity will be used to split the data into two forms in cubble.

@@

In a cubble, there are two forms: nested form and long form, and Figure \ref{fig:illu-cubble} sketches the two forms along with the associated attributes. The decision on which form to use is output-oriented, meaning analysts need to first think about whether the output of a particular operation is identified only by the spatial identifier, or a combination of spatial and temporal identifier. The nested cubble is suitable for working with operations that are only identified by the site and this type of operation can be a pure manipulation of spatial variables, or a summary of temporal variables by site (i.e. the output of counting the number of raining day is only identified by sites and hence should be performed with the nested form). Underneath the nested form, a cubble is built from a row-wise dataframe (\code{rowwise_df}) where each site forms a separate group. This structure simplifies the calculation that involves temporal variables by avoiding the use of \code{map} syntax when working with list-column.

For those operations whose output involves both a spatial and temporal dimension, long form should be used. The long form is identified by both the spatial and temporal identifier and adopts a grouped dataframe (`grouped_df`) to forms each site as a group. Spatial variables are stored separately in a \pkg{tibble} as an special attribute of the long cubble. This design avoids repeating the spatial variables at each time stamp while not dropping information from spatial variables.

@@

## Create a cubble in the nested form

\code{as_cubble()} creates a cubble by supplying the three key components: \code{key} as the spatial identifier; \code{index} as the temporal identifier; and a vector of \code{coords} in the order of longitude first and then latitude. The naming of \code{key} and \code{index} follows the convention in the \pkg{tsibble} package. The cubble created by default is in the nested form. Below is an example of creating a cubble:

@@

There are a few information in the cubble header: the name of the \code{key} variable (`id`) and its number (5), bounding box (`[115.97, -32.94, 133.55, -12.42]`), and also the name of variable nested in the \code{ts} column with its type: \code{date [date], prcp [dbl], tmax [dbl], tmin [dbl]}.

@@

The long cubble is suitable to manipulate the time dimension of the data. The function \code{stretch()} switches the nested cubble into the long cubble by first extracts all the spatial variables into a separate tibble and store in the \code{spatial} attribute and then unnests the \code{ts} column:

```{r echo = TRUE}
cubble_long <- cubble_nested %>% stretch(ts)
cubble_long
```

Notice here that the third line in the header now shows the name and type of spatial variables: \code{lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]}.

@@

Manipulation on the spatial and temporal dimension can be an iterative process. Many times, analysts will need to go back and forth between the nested and long cubble. \code{tamp()} inverses the \code{stretch()} that introduced in the previous section, to switch from a long cubble to the nested cubble:

@@

Some functions require time invariant variables to be in the same table as the time varying ones. In cubble, this can be done through \code{migrate()}, which moves the time invariant variables into the long cubble:

@@

Notice that these added time invariant variables will not be remembered the cubble and will disappear if switched to the nested form and then switched back. Hence they should most likely been used at the last step of the analysis:

@@

In the created nested form, each element of the list-column \code{ts} is of \code{tbl_ts} class so that existing operations on the tsibble class can still be applied:

>>>>

Spatio-temporal data usually come to the analysts in various forms and Figure \ref{fig:illu-input} shows four examples of this. These four examples show the common characters of the spatio-temporal data: 1) a spatial identifier that uniquely identifies each site, 2) a temporal identifier that prescribes the time stamp each site is recorded, 3) Coordinates, comprising of latitude and longitude, that locate the site on the map. These identifiers will be the building blocks for cubble. Another way to structure the spatio-temporal data is to think of it as a data cube with three axes: id, time, and variable. Notice that in the time-variable face of this cube, time invariant variables would share the same value across all the time stamps, while time varying variables have values varying across time. (Maybe add a cube diagram here). This heterogeneity will be used to split the data into two forms in cubble.

@@

In a cubble, there are two forms: nested form and long form, and Figure \ref{fig:illu-cubble} sketches the two forms with the associated attributes. The decision on which form to use in data manipulation is output-oriented, meaning analysts need to first think about whether the output of an operation is identified only by the spatial identifier, or a combination of spatial and temporal identifier. The nested cubble is suitable for working with operations that are only identified by the site and this type of operation can be a pure manipulation of time invariant variables, or an operation that summarises time varying variables into site. Underneath the nested form, a cubble is built from a \code{rowwise_df} data frame where each site forms a separate group. This structure simplifies the calculation that involves temporal variables by avoiding the use of \code{purrr::map()} syntax when working with list-column.

For those operations whose output involves both a spatial and temporal identifier, long form should be used. The long form uses a  \code{`grouped_df`} data frame to forms all the time of a site as a group. Time invariant variables are stored separately as an special attribute of the long cubble. This design avoids repeating the spatial variables at each time stamp while not dropping information from spatial variables.

@@

## Create a cubble in the nested form
A cubble can be created with \code{as_cubble()} by supplying the three key components: \code{key} as the spatial identifier; \code{index} as the temporal identifier; and a vector of \code{coords} in the order of longitude and latitude. The naming of \code{key} and \code{index} follows the convention in the \pkg{tsibble} package. Below is an example of creating a cubble and by default, the cubble created is in the nested form:

@@

There are a few information in the cubble header: the name of the \code{key} variable (`id`) and its number (5), the bounding box: \code{[115.97, -32.94, 133.55, -12.42]} , and also the name of variable nested in the \code{ts} column with its type: \code{date [date], prcp [dbl], tmax [dbl], tmin [dbl]}.

@@

The long cubble is suitable to manipulate the time dimension of the data and  \code{stretch()} switches the nested cubble into the long cubble. It first extracts all the spatial variables into a separate tibble and then unnests the \code{ts} column. The spatial variables are then attached as an attributes of the long cubble and can be accessed through `spatial()`

```{r echo = TRUE}
cubble_long <- cubble_nested %>% stretch(ts)
cubble_long

spatial(cubble_long)
```

Notice that the third line in the header now shows the name and type of spatial variables: \code{lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]}.

@@

Manipulation on the spatial and temporal dimension can be an iterative process and analysts may need to go back and forth between the nested and long cubble. \code{tamp()}, an inverse of \code{stretch()}, switches a long cubble to a nested cubble:

@@

Analysts may need the time invariant/ varying variable to be in the same table to apply certain function. In cubble, this can be done through \code{migrate()}, which moves the time invariant variables into the long cubble:

@@

Notice that these added time invariant variables will not be remembered by cubble and will disappear if switched to the nested form and then switched back. Hence, \code{migrat} should usually be used as the last step of the analysis:

@@

In the created nested form, each element of the list-column \code{ts} is of \code{tbl_ts} class and existing operations on the tsibble class can still be applied:

### 2022-02-07 3-advanced.Rmd
small changes in read-through

Spatial locations can have grouping structure that is either inherent to the data  (i.e. countries are nested in continents), or formed by clustering. Rather than analysing variables in the site level, summarised variables in the cluster level can give a crisper picture of local areas. In cubble, \code{switch_key()} can be used to create a new level of grouping of spatial location by specifying a clustering variable. Figure \ref{fig:illu-hier} illustrates the relationship of cubbles at station and cluster level, in both the long and nested form. By specifying \code{cluster_nested <- station_nested \%>\% switch_key(key = cluster)}, the cubble re-defines the cubble key from \code{id} in \code{station_nested} to \code{cluster} in \code{cluster_nested}. All the spatial variables variant to \code{cluster} are now nested into a \code{.val} column and cluster level variables can be computed in the same fashion as station level variables in \code{station_nested}. Following the same principal of \code{stretch}. the long form cluster level cubble expands the time series variables with both indices: \code{cluster} and \code{id}. All the cluster level variables are stored separately for recovering the nested from from the long form.

@@

\textcolor{red}{refine the beginning.} To be a valid pair of matching, the matched pair from different data sources need to

@@

\code{match_sites()} from \pkg{cubble} matches data based on these two criteria. It first matches the two data sources spatially through computing the pairwise distance based on latitude and longitude. Pairs that pass the spatial matching are then matched temporally through computing the number of matched peaks within a fixed length window. Figure \ref{fig:illu-matching} illustrate this temporal matching in more details. Given two series \code{A} and \code{a}. 3 peaks have been picked in each series. An interval, with default length of 5, is constructed for each peak in series \code{A} and the peaks in series \code{a} are tested against whether they fall into the any of the intervals. In this illustration, there are 2 matches for these two series. Several arguments are available in \code{match_sites()} to fine-tune the matching:

<!-- [can also think about matching as a form of correlation between series from a statistics point of view.] -->

@@

 - \code{spatial_n_keep}. the number of spatial match for each site to keep
 - \code{spatial_dist_max}. the maximum distance allowed for a matched pair 
 - \code{temporal_n_highest}. the number of peak used - 3 in the example above
 - \code{temporal_window}. the length of the interval - 5 in the example above
 - \code{temporal_min_match}. the minimum number of matched peak for a valid matched pair

This functionality can be seen as a matching [@stuart2010matching; @mcintosh2018using] exercise in the spatio-temporal domain to pair up nearby observations. It can also be considered as a form of data fusion [@castanedo2013review; @cocchi2019data], where data from multiple sources are combined together.

@@

Cubble fits in naturally with the interactive graphic pipeline discussed in the literature [@buja1988elements; @buja1996interactive;  @sutherland2000orca; @xie2014reactive; @cheng2016enabling]. Diagram \ref{fig:illu-interactive} illustrates how linking works from the map to the time series with the two forms of a cubble. The map and time series plot is associated with the nested or long cubble respectively and when a user action is captured on the map, the site will be activated in the nested cubble (left). The nested cubble will communicate to the long cubble to activate all the observations with the same \code{id} (middle). The long cubble will then highlight the activated series in the time series plot (right).

@@

Glyph map [@Wickham2012-yr] plots the time series as single glyph on the map and in \proglang{R}, \pkg{GGally} implements the glyph map through the \code{glyphs()} function. The function construct a data frame with calculated position (\code{gx}. \code{gy}. \code{gid}. of each point on the time series using linear algebra (Equation 1 and 2 in @Wickham2012-yr). The data can then be piped into \code{ggplot} to create the glyph map as:

@@

A re-implementation of the glyph map as a ggproto \code{GeomGlyph} has been made in the \pkg{cubble} package so that the glyph map can be created with \code{geom_glyph()}:

@@

Polar glyph map can be specify as a parameter, \code{polar = TRUE}. in the \code{geom_glyph()}. along with \code{width} and \code{height} in either absolute or relative value. Global and local scale can be controlled by the parameter \code{global_rescale}. which default to \code{TRUE} for global scaling. Reference box and line can be added with separate \code{geom_glyph_box()} and \code{geom_glyph_line()}.

>>>>

Spatial locations can have grouping structure either inherent to the data or formed by clustering. Rather than analysing variables in the site level, summarised variables in the cluster level can give a crisper picture of local areas. In cubble, \code{switch_key()} can be used to create a new level of grouping of spatial locations by specifying a clustering variable. Figure \ref{fig:illu-hier} illustrates the relationship of cubbles at station and cluster level, in both the long and nested form. By specifying \code{cluster_nested <- station_nested \%>\% switch_key(key = cluster)}, the cubble re-defines the cubble key from \code{id} in \code{station_nested} to \code{cluster} in \code{cluster_nested}. All the spatial variables variant to \code{cluster} are now nested into a \code{.val} column and cluster level variables can be computed in the same fashion as station level variables in \code{station_nested}.

@@

Cubble provides functionality for matching sites from two data sources. To be a valid pair of matching, the matched pair from different data sources need to:

@@

\code{match_sites()} matches data based on these two criteria: it first matches the two data sources spatially through computing the pairwise distance on latitude and longitude. Pairs that pass the spatial matching are then matched temporally through computing the number of matched peaks within a fixed length window. Figure \ref{fig:illu-matching} illustrates this temporal matching in more details. Given two series \code{A} and \code{a}, 3 peaks have been picked in each series. An interval, with default length of 5, is constructed for each peak in series \code{A} and the peaks in series \code{a} are tested against whether they fall into the any of the intervals. In this illustration, there are 2 matches for these two series. Several arguments are available in \code{match_sites()} to fine-tune the matching:

@@

 - \code{spatial_n_keep}: the number of spatial match for each site to keep
 - \code{spatial_dist_max}: the maximum distance allowed for a matched pair 
 - \code{temporal_n_highest}: the number of peak used - 3 in the example above
 - \code{temporal_window}: the length of the interval - 5 in the example above
 - \code{temporal_min_match}: the minimum number of matched peak for a valid matched pair

This functionality can be seen as a matching exercise [@stuart2010matching; @mcintosh2018using] in the spatio-temporal domain to pair up nearby observations. It can also be considered as a form of data fusion [@castanedo2013review; @cocchi2019data], where data from multiple sources are combined together.

@@

Cubble fits in naturally with the interactive graphic pipeline discussed in the literature [@buja1988elements; @buja1996interactive;  @sutherland2000orca; @xie2014reactive; @cheng2016enabling]. Diagram \ref{fig:illu-interactive} illustrates how linking works from the map to the time series in cubble. The map and time series plot is associated with the nested or long cubble, respectively, and when a user action is captured on the map, the site will be activated in the nested cubble (left). The nested cubble will communicate to the long cubble to activate all the observations with the same \code{id} (middle). The long cubble will then highlight the activated series in the time series plot (right).

@@

Glyph map [@Wickham2012-yr] plots the time series as single glyph on the map and in \proglang{R}, \pkg{GGally} implements the glyph map through the \code{glyphs()} function. The function constructs a data frame with calculated position (\code{gx}. \code{gy}. \code{gid}) of each point on the time series using linear algebra (Equation 1 and 2 in @Wickham2012-yr). The data can then be piped into \code{ggplot} to create the glyph map as:

@@

A re-implementation of the glyph map as a ggproto, \code{GeomGlyph}, has been made in the \pkg{cubble} package and now the glyph map can be created with \code{geom_glyph()}:

@@

Some useful controls over the glyph map is also available in the \code{geom_glyph()} implementation. Polar glyph map can be specify as a parameter, \code{polar = TRUE}. in the \code{geom_glyph()}, along with \code{width} and \code{height} in either absolute or relative value. Global and local scale can be controlled by the parameter \code{global_rescale}. which default to \code{TRUE} for global scaling. Reference box and line can be added with separate \code{geom_glyph_box()} and \code{geom_glyph_line()}.

### 2022-02-07 4-1-basic.Rmd
small changes in read-through

Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world and the dataset \code{weatherdata::historical_tmax} extracts the maximum temperature for `r nrow(weatherdata::historical_tmax)` Australian stations. All the stations start recording from year `r range(weatherdata::historical_tmax$first_year)[2]` with the longest back to year `r range(weatherdata::historical_tmax$first_year)[1]`. The data \code{historical_tmax} is already presented as a cubble, with \code{id} as the key, \code{date} as the index, and \code{c(longitude, latitude)} as the coordinates. This example compares the maximum temperature in two periods: 1971 - 1975 and 2016 - 2020 for stations in Victoria and New South Wales.

@@

There are a few stations which don't have records during 1971 - 1975 and further investigation shows that while the first and last year of each series is recorded, the missing years in this period is not reported. These stations are filtered out by examining whether the summarised time series has a total of 24 months. The long cubble needs to be switched to the nested form for this spatial operation using \code{tamp()}:

>>>>

Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world. The dataset \code{weatherdata::historical_tmax} extracts the maximum temperature for `r nrow(weatherdata::historical_tmax)` Australian stations from GHCN with starting from year `r range(weatherdata::historical_tmax$first_year)[2]`. \code{weatherdata::historical_tmax} is already in a cubble, with \code{id} as the key, \code{date} as the index, and \code{c(longitude, latitude)} as the coordinates. This example compares the maximum temperature in two periods: 1971 - 1975 and 2016 - 2020 for stations in Victoria and New South Wales.

@@

A few stations don't have records during 1971 - 1975 and further investigation shows that while the first and last year of each series is recorded, the missing years in this period is not reported. These stations are filtered out by examining whether the summarised time series has a total of 24 months. The long cubble needs to be switched to the nested form for this spatial operation using \code{tamp()}:

### 2022-02-07 4-2-agg.Rmd
small changes in read-through

In the previous example, there has already been some overlapping of the glyphs for a few stations near (151E, 34S) and (152E, 33S) and this will be a problem when mapping more stations in the national level. Aggregation can be helpful to group series into clusters before visualising the cluster with glyph map.

This example shows how to organise data at both level with \code{switch_key()}. The dataset used is \code{weatherdata::climate_full}, which records daily precipitation and maximum/ minimum temperature for `r nrow(weatherdata::climate_full)` stations in Australia from 2016 to 2020. A simple kmean algorithm based on the distance matrix is used here to create 20 clusters. This creates \code{station_nested} as a station level nested cubble with a cluster column indicating the group each station belongs to. More complex algorithms can be used for other problem, as long as there is a mapping from each station to a cluster.

@@

Long form cubble at both levels can be access through stretching the nested form and operations applicable to long form cubble are still available. With access to both station and cluster level cubbles, various plots can be made to understand the cluster. Figure \ref{fig:basic-agg} shows two example plots that can be made with this data: subplot A makes a glyph map with cluster level cubble and subplot B inspects the station membership of each cluster using the nested station level cubble.

>>>>

In the previous example, there has already been some overlapping of the glyphs for a few stations near (151E, 34S) and (152E, 33S) and this will be a problem when mapping more stations in the national level. Aggregation can be helpful to group series into clusters before visualising the cluster with glyph map. This example shows how to organise data at both level with \code{switch_key()}.

\code{weatherdata::climate_full}, also extracted from GHCN, records daily precipitation and maximum/ minimum temperature for `r nrow(weatherdata::climate_full)` stations in Australia from 2016 to 2020. A simple kmean algorithm based on the distance matrix is used to create 20 clusters. This creates \code{station_nested} as a station level nested cubble with a cluster column indicating the group each station belongs to. More complex algorithms can be used for other problem, as long as there is a mapping from each station to a cluster.

@@

Long form cubble at both levels can be accessed through stretching the nested form and with access to both station and cluster level cubbles, various plots can be made to understand the cluster. Figure \ref{fig:basic-agg} shows two example plots that can be made with this data: subplot A is a glyph map made with the cluster level cubble in the long form and subplot B inspects the station membership of each cluster using the station level cubble in the nested form.

### 2022-02-07 4-3-water-level.Rmd
small changes in read-through

Bureau of Meteorology also collects [water data](http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full) from river gauges, which includes electrical conductivity, turbidity, water course discharge, water course level, and water temperature. In particular, water level will interactive with precipitation from the climate data since rainfall will raise the water level in the river. Figure \ref{fig:matching-map} gives the location of available weather station and water gauges in Victoria.

@@

From the map, a few water gauges and weather stations are close to each other and there is the potential that fluctuation of the water level can reflect the precipitation measured by the climate station. As introduced in Section 3.2, \code{match_sites()} can be used to match one source of data with another source in a cubble. Here \code{prcp} in \code{climate} will be matched to \code{Water_course_level} in \code{river}, which is specified in the argument \code{temporal_var_to_match}. \code{temporal_independent} controls the dataset that will be used to construct the interval (see the illustration in section 3.2 for details on this). Here the goal is to see if precipitation will be reflected by the water level in the river and hence a lagged effect of water level on precipitation. This will put precipitation data, \code{climate}, as the independent. Given there is one year worth of data, the number of peak (\code{temporal_n_highest}) to consider is slightly raised from a default 20 to 30. \code{temporal_min_match} filters out pairs don't have enough match and to return all the pairs, set \code{temporal_min_match} to 0.

@@

The output from matching is also a cubble, with additional column \code{dist} and \code{group} produced from spatial matching and \code{n_match} from the temporal matching.

>>>>

Bureau of Meteorology collects [water data](http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full) from river gauges and this includes variables: electrical conductivity, turbidity, water course discharge, water course level, and water temperature. In particular, water level will interactive with precipitation from the climate data since rainfall will raise the water level in the river. Figure \ref{fig:matching-map} gives the location of available weather station and water gauges in Victoria.

@@

From the map, a few water gauges and weather stations are close to each other and the fluctuation of the water level could be matched up with precipitation measured by the climate station. As introduced in Section 3.2, \code{match_sites()} can be used to match one source of data with another source in a cubble. Here \code{Water_course_level} in \code{river} will be matched to \code{prcp} in \code{climate} in 2020. The two datasets need to be specified as the first two arguments and the variable to match can be specified in \code{temporal_var_to_match} using the \code{by} syntax in \code{join}. \code{temporal_independent} controls the variable used to construct the interval and the goal here is to see if precipitation will be reflected by the water level in the river. This puts precipitation \code{prcp}, as the independent. Given there is one year worth of data, the number of peak (\code{temporal_n_highest}) to consider is slightly raised from a default 20 to 30 and \code{temporal_min_match} is raised correspondingly. To return all the pairs of the match, \code{temporal_min_match} can be set to 0.

@@

The output from matching is also a cubble, with additional column \code{dist} and \code{group} produced from spatial matching and \code{n_match} from temporal matching.

### 2022-02-07 4-4-netcdf.Rmd
small changes in read-through

NetCDF (Network Common Data Form) is a data format commonly used in climatology community to deliver global mapping of atmosphere, ocean, and land. It has two major components: dimension to define the spatio-temporal grid (longitude, latitude, and time) and variable which populates the defined grid. Attributes are usually associated with dimension and variable in the NetCDF format data and a [metadata convention for climate and forecast](http://cfconventions.org/) has been designed to standardise this data format. Current packages on manipulating NetCDF data in R include a high-level R interface: \pkg{ncdf4} [@ncdf4], a low-level interface that calls C-interface: \pkg{RNetCDF} [@rnetcdf; @michna2013rnetcdf], and a tidyverse implementation: \pkg{tidync} [@tidync].

Cubble provides an \code{as_cubble()} method to coerce the \code{ncdf4} class from the \pkg{ncdf4} package into a \code{cubble}. It maps each combination of longitude and latitude into an \code{id} (the \code{key}) and nests time and variables in the nested form:

@@

Memory limit with NetCDF data in cubble depends on longitude grid point x latitude grid point x time grid point x number of variable. Cubble can handle slightly more than 300 x 300 (longitude x longitude) grid points for 3 variables in one year and spatial grid can be reduced to trade for longer time period and more variables. A 300 by 300 spatial grid can be a bounding box of [100, -80, 180, 0] at 0.25 degree resolution or global bounding box [-180, -90, 180, -90] at 1 degree resolution. Subsetting longitude and latitude grid is available through \code{long_range} and \code{lat_range} if the NetCDF file has finer resolution than needed.

@@

Figure \ref{fig:netcdf} gives an example of reproducing Figure 19 in @hersbach2020era5 using ERA5 data. ERA5 data [@hersbach2020era5] is the latest reanalysis of global atmosphere, land surface, and ocean waves from 1950 onwards and is available in the NetCDF format. It can be directly downloaded from the [Copernicus Climate Data Store (CDS)](https://cds.climate.copernicus.eu/cdsapp#!/dataset/reanalysis-era5-pressure-levels?tab=overview) website or programmatically via R package \pkg{ecmwfr} [@ecwmfr]. Variable Specific humidity and geopotential are queried on the 10 hPa pressure level for four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04 and once downloaded, the data can be read into a cubble using the code above. Readers who are interested in the details of can refer to @simmons2020global and @simmons2005ecmwf.

>>>>

NetCDF (Network Common Data Form) is a data format commonly used in the climatology community to deliver global mapping of atmosphere, ocean, and land. It has two major components: *dimension* for defining the spatio-temporal grid (longitude, latitude, and time) and *variable* that populates the defined grid. Attributes are usually associated with dimension and variable in the NetCDF format data and a [metadata convention for climate and forecast](http://cfconventions.org/) has been designed to standardise the format of the attributes. Current packages for manipulating NetCDF data in R include a high-level R interface: \pkg{ncdf4} [@ncdf4], a low-level interface that calls C-interface: \pkg{RNetCDF} [@rnetcdf; @michna2013rnetcdf], and a tidyverse implementation: \pkg{tidync} [@tidync].

Cubble provides an \code{as_cubble()} method to coerce the \code{ncdf4} class from the \pkg{ncdf4} package into a \code{cubble}. It maps each combination of longitude and latitude into an \code{id} as the \code{key}:

@@

Memory limit with NetCDF data in cubble depends on longitude grid point x latitude grid point x time grid point x number of variable. Cubble can handle slightly more than 300 x 300 (longitude x longitude) grid points for 3 variables in one year and spatial grid can be reduced to trade for longer time period and more variables. A 300 by 300 spatial grid can be a bounding box of [100, -80, 180, 0] at 0.25 degree resolution or a global bounding box [-180, -90, 180, -90] at 1 degree resolution. Subsetting longitude and latitude grid is available through \code{long_range} and \code{lat_range} if the NetCDF file has finer resolution than needed.

@@

Figure \ref{fig:netcdf} gives an example of reproducing Figure 19 in @hersbach2020era5 using ERA5 data. ERA5 data [@hersbach2020era5] is the latest reanalysis of global atmosphere, land surface, and ocean waves from 1950 onwards and is available in the NetCDF format from European Centre for Medium-Range Weather Forecasts (ECMWF). The data can be directly downloaded from [Copernicus Climate Data Store (CDS)](https://cds.climate.copernicus.eu/cdsapp#!/dataset/reanalysis-era5-pressure-levels?tab=overview) website or programmatically via an R package \pkg{ecmwfr} [@ecwmfr]. Variable *specific humidity* and *geopotential* are queried on the 10 hPa pressure level on four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04. Once downloaded, the data can be read into a cubble as:

```{r eval = FALSE, echo = TRUE}
raw <- ncdf4::nc_open(here::here("data/era5-pressure.nc"))
dt <- as_cubble(raw, vars = c("q", "z"))
```

Readers who are interested in the analysis of the figure can refer to @simmons2020global and @simmons2005ecmwf for more details.
